Evaluating impacts of Density Dependent processes on Blue whiting via MSE simulations

### 2023-03-15 Rmd/report whb.Rmd
Updated WHB + report; Second GAMM

@@

Estimating density-dependence from observations was not really possible for `r mystkname`. Therefore, density dependence in mass at age, maturity at age and natural mortality has only be explored from theoretical considerations in this report.

@@

**GAMM-at-age**

@@

**Linear model of density dependence in mass-at-age**

The change in relative mass-at-age (i.e. observed - mean values) is plotted against total biomass, this allows the relationship between growth and total biomass to be compared across all ages. Here, we plotted both the GAMM model and the Linear Model variants. It is seen that there is still some weak information on density dependence in the Linear Model (slope = beta = -0.11). A summary of the linear model is in the table below.

@@

fig_nums(
  name    = "cond_lmwaa", level = 1, display = FALSE,
  caption = paste0(mystkname, ". LM analysis of density dependence and mass-at-age."))

knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk, "lm_waa.jpg", sep="_")))

@@

*`r fig_nums("cond_lmwaa")`*

@@

tab_nums(
  name    = "cond_lmsummary", level = 1, display = FALSE,
  caption = paste0(mystkname, ". LM summary of density dependence and mass-at-age."))

cat(readLines(file.path(tablesdir, paste(section,mystk, "lm summary.txt", sep="_"))), sep = '\n')

@@

*`r tab_nums("cond_lmsummary")`*

@@

**@Laurie: why would DD in natural mortality have such a minimal impact? Did you see that the M levels overall are lower than the assumed M in the assessment (=0.2)**

@@

Estimating density-dependence from observations was only possible for mass-at-age but not maturity-at-age as the latter were assumed by ICES not to have varied over time. However, assuming that maturity is a function of mass-ata-ge allowed density deoendence to be modeeled. Natural mortality was assumed to be time and age invariant by ICES (i.e. 0.2), therefore density dependence in natural mortality can only be explored from theoretical considerations.

>>>>

**Laurie: why would DD in natural mortality have such a minimal impact? Did you see that the M levels overall are lower than the assumed M in the assessment (=0.2)**

@@

Estimating density-dependence from observations was not really possible for `r mystkname` because there was no clear relationship between mass-at-age and total biomass. When a second GAMM model was fitted with weight in relation to total biomass, year and age, we found that the mass-at-age is better described by a year effect than a biomass effect. In order to still explore the impacts of density dependence on the performance of the MSE, we used the density dependent factor for Atlantic mackerel (beta=-0.18) to parameterize the weight-biomass relationship.

@@

**GAMM-at-age: biomass only**

@@

**GAMM-at-age: biomass and year smoothers**

Because no clear relationship between mass-at-age and total biomass was established, we fitted a second GAMM model with weight in relation to total biomass, year and age. This demonstrated that the mass-at-age is better described by the year effect than the biomass effect (lower p-value).

In order to still explore the impacts of density dependence on the performance of the MSE, we used the density dependent factor for Atlantic mackerel (beta=-0.18) to parameterize the following relationship:

@@

fig_nums(
  name    = "cond_gammwaa2", level = 1, display = FALSE,
  caption = paste0(mystkname, ". GAMM analysis of mass-at-age against biomass, year and age"))

knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk, "gamm2.jpg", sep="_")))

@@

*`r fig_nums("cond_gammwaa2")`*

@@

tab_nums(
  name    = "cond_gammsummary2", level = 1, display = FALSE,
  caption = paste0(mystkname, ". GAMM summary of mass-at-age against biomass, year and age"))

cat(readLines(file.path(tablesdir, paste(section,mystk, "gamm2 summary.txt", sep="_"))), sep = '\n')

@@

*`r tab_nums("cond_gammsummary2")`*

@@

@@

Estimating density-dependence from observations was not really possible for `r mystkname` because there was no clear relationship between mass-at-age and total biomass. When a second GAMM model was fitted with weight in relation to total biomass, year and age, we found that the mass-at-age is better described by a year effect than a biomass effect. In order to still explore the impacts of density dependence on the performance of the MSE, we used the density dependent factor for Atlantic mackerel (beta=-0.18) to parameterize the weight-biomass relationship.
